---
title: "DegreeDist"
author: "Emily Pollock"
date: "7/26/2019"
output: html_document
---

```{r setup, include=FALSE}
options("scipen" = 10)

knitr::opts_chunk$set(echo = TRUE)
library(ergm.ego)
library(tidyverse)
library(reshape2)
library(srvyr)
library(plotly)

### import objects ###

### make egodata object ###
alters <- readRDS("~/NSFG_DATA/Objects/alters.rds")
egos <- readRDS("~/NSFG_DATA/Objects/egos.rds")
# make subsets to test  
egosF <- egos %>% filter(sex %in% "F")
egosH <- egos %>% filter(race %in% "h")

dat <- egodata(egos=egos, alters=alters, egoIDcol = "ego")
dat_F <- egodata(egos=egosF, alters=alters, egoIDcol = "ego")
dat_H <- egodata(egos=egosH, alters=alters, egoIDcol = "ego")

### wide data, weighted and unweighted ###
nsfg <- readRDS("~/NSFG_DATA/Full/nsfg_complete.rds")
```

## Basic Problem 

Degree Distribution by sex and other attributes is different when calculated using ergm.go or when using summary stats on "deg.main" in original wide data format. 

ergm.ego uses a different data format which has two components:  
1) alter list of all ACTIVE partners, the ego id they correspond to, and demographic info  
2) ego list of ALL EGOS in dataset, ego id, and demographic and other descriptive vars  

original data is formated with 1 row per ego and all alter / other info are column variables 

### unweighted degree dist (prop & frequencies) from ergm.ego  
```{r ergmego, echo=FALSE}
deg <- degreedist(dat, by="sex", plot=F, freq = F); round(deg, 3)
degFreq <- degreedist(dat, by="sex", plot=F, freq = T); degFreq
```


### unweighted degree dist from original data  
("deg.main" = sum of active partnerships)
```{r widedeg, echo=FALSE}
deg_wide <- nsfg %>% 
  group_by(sex, deg.main) %>% 
  count() %>% 
  group_by(sex) %>%
  mutate(prop = n/sum(n)) %>%
  mutate(prop = round(prop, 3))

prop <- deg_wide %>%
  dcast(sex~deg.main, value.var = "prop")

freq <- deg_wide %>%
  dcast(sex~deg.main, value.var = "n")
prop; freq
```

## Basic Troubleshooting 

### Pulling data from same number of egos? -- yes

egodata:  
```{r rowSums-datun}
rowSums(degFreq)
```

original:
```{r rowSums-nsfg}
rowSums(freq[,2:5])
```

### Same number of active alters? -- yes 

egodata: 
```{r rows-alters}
nrow(dat$alters)
```

original: 
```{r sum-degmain}
sum(nsfg$deg.main)
```

### Are the deg dists same if you don't stratify by sex? -- yes 

egodata:
```{r deg-all}
degreedist(dat, plot=F, freq=T)
```

original:  
```{r degdist-all}
nsfg %>% count(deg.main)
```

### Checking that stratifying by other vars also produces different results - yes
```{r deg-race}
degreedist(dat, by="race", plot=F, freq=T)
```

original:  
```{r degdist-racr}
nsfg %>% group_by(race) %>% count(deg.main) %>% dcast(race~deg.main, value.var = "n")
```
neither method drops any obvervations, working with same # of egos/alters. 

ALSO colSums are equal, so they agree how many deg0, deg1, etc exists, but they don't agree on say how many females are deg0

what in the hell

Something in the degreedist function apportions degree differently when stratifying by a variable than when grouping and summarizing over the deg.main variable. 

## subset - females -- THIS WORKS 

egodata:  
```{r sub}
degsub <- degreedist(dat_F, plot=F, freq=T); degsub
degsubRace <- degreedist(dat_F, by="race", plot=F, freq=T); degsubRace
```

original:  
```{r newdata}
nsfg %>% filter(sex %in% "F") %>% count(deg.main)

nsfg %>% filter(sex %in% "F") %>% group_by(race) %>% count(deg.main) %>% dcast(race~deg.main)
```


## subset - hispanics -- by sex doesn't work 
```{r sub2}
degreedist(dat_H, plot=F, freq = T)

nsfg %>% filter(race %in% "h") %>% count(deg.main)

degreedist(dat_H, by="sex", plot=F, freq=T)

nsfg %>% filter(race %in% "h") %>% group_by(sex) %>% count(deg.main)
```


# degree dist by deg.main

```{r degmain}
degreedist(dat, by="deg.main", freq = T, plot=F)
```

```{r deg3}
tb <- table(dat$alters[["ego"]])
deg3 <- tb[which(tb==3)]
egosTest <- as.numeric(names(deg3))

ego <- nsfg %>% filter(ego %in% egosTest) 
```

```{r pavel-function, echo=FALSE}

degreedist.egodata <- function(object, freq = FALSE, prob = !freq, 
                               by = NULL, brgmod = FALSE, main = NULL, plot = TRUE, weight = TRUE, ...){
  egodata <- object
  if(!weight) egodata$egoWt[] <- 1
  
  color <- "#83B6E1"
  beside <- TRUE
  
  ylabel <- if(prob) "Proportion" else "Frequency"
  if(!is.null(by)) ylabel <- paste(ylabel, "(within attr level)")
  
  egoIDcol <- egodata$egoIDcol
  
  egodata$egos[[egoIDcol]] <- factor(egodata$egos[[egoIDcol]])
  egodata$alters[[egoIDcol]] <- factor(egodata$alters[[egoIDcol]], levels=levels(egodata$egos[[egoIDcol]])) 
  egodata$egos[[egoIDcol]] <- as.integer(egodata$egos[[egoIDcol]])
  egodata$alters[[egoIDcol]] <- as.integer(egodata$alters[[egoIDcol]])
  
  degtable <- rep(0, nrow(egodata$egos))
  degtable[as.numeric(names(table(egodata$alters[egoIDcol])))] <- table(egodata$alters[egoIDcol])
  
  if(is.null(by)){
    deg.ego <- xtabs(egodata$egoWt~degtable)
    names(dimnames(deg.ego)) <- "degree"
    degrees <- as.integer(names(deg.ego))
  }else{
    deg.ego <- xtabs(egodata$egoWt~egodata$egos[[by]]+degtable)
    names(dimnames(deg.ego)) <- c(by, "degree")
    levs <- rownames(deg.ego)
    degrees <- as.integer(colnames(deg.ego))
    ncolors <- dim(deg.ego)[1]
    if(ncolors == 2){
      color <- c("#eff3ff", "#377FBC")
    } else if(ncolors < 10){
      color <- RColorBrewer::brewer.pal(ncolors,"Blues")
    } else if(ncolors >= 10){
      color <- grDevices::colorRampPalette(RColorBrewer::brewer.pal(9,"Blues"))(ncolors)
    }
    
    ltext <- levs
    lfill <- c(color, 0)
    ltitle <- by
    maxfreq <- max(colSums(deg.ego))
  }
  if(prob){
    if(is.null(by)){
      scaledeg <- sum(deg.ego)
      deg.ego <- deg.ego/scaledeg
      maxfreq <- max(deg.ego, na.rm = TRUE)
    } else {
      scaledeg <- rowSums(deg.ego)
      deg.ego <- deg.ego/scaledeg
      deg.ego <- deg.ego
      maxfreq <- max(max(deg.ego, na.rm = TRUE))
      beside <- TRUE
    }
  }
  
  if(plot){
    if(brgmod) {
      ppopsize.mul <- max(egodata$egoWt)/min(egodata$egoWt)
      brgdraws <- simulate(suppressMessages(ergm.ego(egodata ~ edges, control=control.ergm.ego(ppopsize=nrow(egodata$egos)*ppopsize.mul))), nsim = 50, ...)
      deg.brg <- summary(brgdraws ~ degree(degrees))/ppopsize.mul
      brgmeans <- apply(deg.brg, MARGIN = 2, FUN = mean)
      brgsd <- apply(deg.brg, MARGIN = 2, FUN = sd)
      upper <- brgmeans + 2 * brgsd
      lower <- brgmeans - 2 * brgsd
      
      if(prob){
        if(is.null(by)){
          brgmeans <- brgmeans/scaledeg
          upper <- upper/scaledeg
          lower <- lower/scaledeg
        } else {
          upper <- upper/sum(brgmeans)
          lower <- lower/sum(brgmeans)
          brgmeans <- brgmeans/sum(brgmeans)
        }
        
      }
      maxfreq <- max(maxfreq, upper, na.rm = TRUE)
    }
    
    baraxis <- barplot(deg.ego, xlab = "Degree", ylab = ylabel,
                       col = color, beside = beside, plot = TRUE,
                       ylim = c(0, maxfreq), main = main)
    
    if(brgmod){
      baraxis <- if(is.null(by)){
        baraxis - 0.15
      } else {
        colMeans(baraxis)
      }
      points(x = baraxis, y = brgmeans, col = "firebrick",
             lwd = 1, pch = 18, cex = 1.25)
      suppressWarnings(arrows(x0 = baraxis, y0 = upper,
                              x1 = baraxis, y1 = lower,
                              code = 3, length = 0.1, 
                              angle = 90, col = "firebrick"))
    } 
    if(!is.null(by)){
      legend(x="top", legend = ltext, title = ltitle, fill = lfill, bg="white")
    }
  }
  
  if(plot) invisible(deg.ego) else deg.ego
}
```