# Alter Race Issues {#missingrace}

## Problem
__alter race was not asked in a variety of situations__

We HAVE alter race for all active partnerships. 
For MALE EGOS, we HAVE race for all previous cohabs and marriages. 
We also have a few one-time partner races for MALES if the one-time partner was their most recent partner. 

We DO NOT HAVE alter race for any inactive partnerships or one-time partnerships for females. 
We DO NOT HAVE alter race for inactive "others" for males. 
We DO NOT HAVE alter race for male one-time partners that were not their most recent partner. 

```{r race-misings, echo=FALSE}

suppressMessages(library(tidyverse))
suppressMessages(library(srvyr))
suppressMessages(library(kableExtra))
suppressMessages(library(reshape2))
#suppressMessages(library(plotly))
library(ergm.ego)
library(ggplot2)

# full dataset (egos < 30)
nsfg <- readRDS("~/NSFG_DATA/Full/nsfg_complete.rds")
svy <- as_survey_design(nsfg, weights = weight, ids = ego)

# long alters list 
alters <- readRDS("~/NSFG_DATA/Objects/altersegos_survdat.rds")
svy_alters <- as_survey_design(alters, weights = weight, ids = ego)


missingsActive <- alters %>% filter(active ==1) %>% group_by(e.sex, reltype, race) %>% count() %>% dcast(e.sex+reltype~race, value.var = "n")

missingsInactive <- alters %>% filter(active ==0) %>% group_by(e.sex, reltype, race) %>% count() %>% dcast(e.sex+reltype~race, value.var = "n")

MaleOtherInactives <-  alters %>% filter(active ==0, e.sex %in% "M", reltype %in% "Other") %>% 
  group_by(e.intyr, race) %>% count() %>% dcast(e.intyr~race, value.var = "n")


missingsOneTimes <- alters %>% filter(once==1 & active==0) %>% group_by(e.sex, reltype, race, .drop=FALSE) %>% count() %>% dcast(e.sex+reltype~race, value.var = "n")

OTmales <- alters %>% filter(once==1 & active==0, e.sex %in% "M") %>% group_by(e.intyr, alter, race, .drop=FALSE) %>% count() %>% dcast(e.intyr+alter~race, value.var = "n")


```

## Active Partnerships
`r kable(missingsActive)` 

## Inactive Partnerships
`r kable(missingsInactive)` 

### close up on "other" for males  
switched how they asked the question? basically all miss/ref/dk and n.asked probably missing for same reason

`r kable(MaleOtherInactives)`

### what about the male one-times that we know about? 


## One-Times
`r kable(missingsOneTimes)` 

for males, by alter 
`r kable(OTmales)` 

## Imputation 

proposal: impute partner race for missngs 

Males:   
 - impute missing one-times based on most recent one-times 
 
 Females:  
  - impute missings one-times based on transposed male one-time responses 
  
  
  other questions - 
  are female reports of mixing different by cohab/marriage versus other? - mostly yea
  are male reports of active cohab/marriage different than inactive marriage/cohab? - also yes a bit 

```{r more-mixing, echo=FALSE, include=FALSE}
# females - active rels by marriage/cohab and other 
# import objects
alters <- readRDS("~/NSFG_DATA/Objects/alters.rds")
egos <- readRDS("~/NSFG_DATA/Objects/egos.rds")
egosF <- egos %>% filter(sex %in% "F")
alters_marcoh <- alters %>% filter(optype %in% 1:4)
alters_other <- alters %>% filter(optype ==5)
# make ego data object
dat <- egodata(egos=egos, alters=alters, egoWt = egos$weight, egoIDcol = "ego")
datFmarcoh <- egodata(egos=egosF, alters=alters_marcoh, egoWt = egos$weight, egoIDcol = "ego")
datFother <- egodata(egos=egosF, alters=alters_other, egoWt = egos$weight, egoIDcol = "ego")

Fmarcoh <- round(mixingmatrix(datFmarcoh, "race", rowprob = T), 3)
Fmarcoh <- Fmarcoh[c(1,2,5,6), c(1,2,5,6)]
Fmarcoh <- melt(Fmarcoh) 

Fmarcoh %>% ggplot(aes(ego, alter)) + 
  geom_point(color="blue", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.02)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Race Mixing, Female Active MarCoh")

Fother <- round(mixingmatrix(datFother, "race", rowprob = T), 3)
Fother <- Fother[c(1,2,4,5), c(1,2,4,5)]
Fother <- melt(Fother) 

Fother %>% ggplot(aes(ego, alter)) + 
  geom_point(color="blue", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.02)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Race Mixing, Female Active Other")
```

```{r even-more-mixing, echo=FALSE, include=FALSE}

# objects
egosM <- egos %>% filter(sex %in% "M")
alters_active_marcoh <- alters %>% filter(optype %in% 1:4)

# need survdat object b/c has inactives 
alters_all <- readRDS("~/NSFG_DATA/Objects/altersegos_survdat.rds")
alters_inactive_marcoh <- alters_all %>% filter(active==0, optype %in% 1:4)

# make ego data object
datMmarcohA <- egodata(egos=egosM, alters=alters_active_marcoh , egoWt = egos$weight, egoIDcol = "ego")
datMmarcohI <- egodata(egos=egosM, alters=alters_inactive_marcoh, egoWt = egos$weight, egoIDcol = "ego")

MmarcohA <- round(mixingmatrix(datMmarcohA, "race", rowprob = T), 3)
MmarcohA  <- MmarcohA [c(1,2,4,5), c(1,2,4,5)]
MmarcohA  <- melt(MmarcohA ) 

MmarcohA  %>% ggplot(aes(ego, alter)) + 
  geom_point(color="blue", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.02)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Race Mixing, Male Active MarCoh")

MmarcohI <- round(mixingmatrix(datMmarcohI, "race", rowprob = T), 3)
MmarcohI <- MmarcohI[c(1,2,4,5), c(1,2,4,5)]
MmarcohI <- melt(MmarcohI) 

MmarcohI %>% ggplot(aes(ego, alter)) + 
  geom_point(color="blue", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.02)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Race Mixing, Male Inactive MarCoh")

```


  
  
## Imputation Targets

need: 
Male egos:
  - active "other" alter race prob given male ego race
  - most recent "one-time" alter race given male ego race 
 
 Females:  
  - active "cohab/mar" alter race prob given female ego race  
  - active "other" alter race prob given female ego race  
  - transposed probs of male ego most recent one-time dist 
  
### Male Active Others
```{r male-target-probs, echo=FALSE}
# set up alters_other to exclude the few missings
alters_other <- alters_other %>% mutate(race = ifelse(race %in% "miss/ref/dk", NA,
                                                      ifelse(race %in% "n.ask", NA, race)))

MaleOthers <- egodata(egos=egosM, alters=alters_other, egoWt = egos$weight, egoIDcol = "ego")

MaleActiveOthers <- round(mixingmatrix(MaleOthers, "race", rowprob = F),0)
MaleActiveOthersProbs <- round(mixingmatrix(MaleOthers, "race", rowprob = T),3)

# recent one time alters
alters_once <- alters_all %>% filter(active==0, once==1, alter==1) %>%  
                              mutate(race = ifelse(race %in% "miss/ref/dk", NA,
                                                      ifelse(race %in% "n.ask", NA, race)))

MaleRecentOnce <- egodata(egos=egosM, alters=alters_once, egoWt = egos$weight, egoIDcol = "ego")

onces <- round(mixingmatrix(MaleRecentOnce, "race", rowprob = F),0)
oncesProb <- round(mixingmatrix(MaleRecentOnce, "race", rowprob = T),3)

```

`r kable(MaleActiveOthersProbs)`

### Male Recent One-Times

`r kable(oncesProb)`

### Female Active Mar/Coh
```{r fem-targets, echo=FALSE}
alters_marcoh <- alters %>% filter(optype %in% 1:4) %>%
                            mutate(race = ifelse(race %in% "miss/ref/dk", NA,
                                                      ifelse(race %in% "n.ask", NA, race)))

alters_other <- alters %>% filter(optype ==5) %>%
                            mutate(race = ifelse(race %in% "miss/ref/dk", NA,
                                                      ifelse(race %in% "n.ask", NA, race)))

FemaleMarcoh <- egodata(egos=egosF, alters=alters_marcoh, egoWt = egos$weight, egoIDcol = "ego")
FemaleOther <- egodata(egos=egosF, alters=alters_other, egoWt = egos$weight, egoIDcol = "ego")

Marcoh <- round(mixingmatrix(FemaleMarcoh, "race", rowprob = F),0)
MarcohProb<- round(mixingmatrix(FemaleMarcoh, "race", rowprob = T),3)

Other <- round(mixingmatrix(FemaleOther, "race", rowprob = F),0)
OtherProb<- round(mixingmatrix(FemaleOther, "race", rowprob = T),3)
```

`r kable(MarcohProb)`

### Female Active Other
`r kable(OtherProb)`

### Female One-Times
```{r transposed-ots, echo=FALSE}
Fonces <- t(onces)
FoncesProb <- round(Fonces/rowSums(Fonces),3) 
```

`r kable(FoncesProb)`

