# Concurrency & Racial Assortativity 

CURRENTLY - CROSS SECTIONAL CONCURRENCY 

thinking about concurrency and racial assortativity 

previous work shows concurrency drives transmission, and racial asssortativity can maintain / exacerbate disparities in prevalence 

but what about assortativity by concurrency levels? 
is assortativity similar between monogramous people and those who have concurrent partners? 
what about among people who have concurrent parters? 

this section takes a look at what is observed in NSFG - to see if there are difference in assortativity and if this idea is worth exploring (and likely simulating to understand the effects)

WILL NEED TO IMPUTE PARTNER RACE FOR MOST NON-ACTIVE PARTNERSHIPS IF WE WANT TO USE A YEAR-SUMMARY MEASURE 



```{r active-alters, echo=FALSE}
library(tidyverse)
library(reshape2)
library(srvyr)
# import
dat <- readRDS("~/nsfg_analysis/data/alters_egos.rds")

## limit to active partnerships 
dat <- dat %>% filter(active==1)

## new variables
# t_o = relationship dur at 1 year prior to interview = edge_age_month-12, if edge_age_month-12 is negative, 0
# t_c = relationship month end (time = 1-12), if still active, 12
dat$t_o <- ifelse((dat$edge_age_month - 12 >= 0), dat$edge_age_month-12, 0)
dat$t_c <- dat$edge_age_month

dat$t_o_years <- dat$t_o/12
dat$t_c_years <- dat$t_c/12
```


see *insert chapter*  for deg distribution

here: what is the race mixing matrix for egos who report only 1 partner
then: what is the race mixing matrix for egos who report 2 partners
then: what is the race mixing matrix for egos who report 3 partners

## Unweighted Mixing Plots 
```{r mixing-unw, echo=FALSE}
dat <-  dat %>% filter(race %in% "b" | race %in% "h"| race %in% "o" |race %in% "w")

monog <- dat %>% 
          filter(e.deg.active ==1) %>%
          group_by(e.sex, e.race) %>%
          count(race) %>%
          mutate(prop = n/sum(n))

g1 <- ggplot(monog, aes(x=e.race, y=race)) + 
  geom_point(color="blue", alpha=0.3, aes(size=prop)) +
  geom_text(aes(label=round(prop,2))) +
  scale_size_area(max_size = 30) +
  coord_flip() +
  theme(legend.position="none") + 
  labs(title = "Race Mixing Among Egos w/ 1 Active Partner")

g1 + facet_grid(cols=vars(e.sex))


conc <- dat %>% 
          filter(e.deg.active > 1) %>%
          group_by(e.sex, e.race) %>%
          count(race) %>%
          mutate(prop = n/sum(n))

g_conc <- ggplot(conc, aes(x=e.race, y=race)) + 
  geom_point(color="blue", alpha=0.3, aes(size=prop)) +
  geom_text(aes(label=round(prop,2))) +
  scale_size_area(max_size = 30) +
  coord_flip() +
  theme(legend.position="none") + 
  labs(title = "Race Mixing Among Egos w/ Concurrent")

g_conc + facet_grid(cols=vars(e.sex))

conc2 <- dat %>% 
          filter(e.deg.active ==2) %>%
          group_by(e.sex, e.race) %>%
          count(race) %>%
          mutate(prop = n/sum(n))

g2 <- ggplot(conc2, aes(x=e.race, y=race)) + 
  geom_point(color="blue", alpha=0.3, aes(size=prop)) +
  geom_text(aes(label=round(prop,2))) +
  scale_size_area(max_size = 30) +
  coord_flip() +
  theme(legend.position="none") + 
  labs(title = "Race Mixing Among Egos w/ 2 Active Partners")

g2 + facet_grid(cols=vars(e.sex))

conc3 <- dat %>% 
          filter(e.deg.active ==3) %>%
          group_by(e.sex, e.race) %>%
          count(race) %>%
          mutate(prop = n/sum(n))

g3 <- ggplot(conc3, aes(x=e.race, y=race)) + 
  geom_point(color="blue", alpha=0.3, aes(size=prop)) +
  geom_text(aes(label=round(prop,2))) +
  scale_size_area(max_size = 30) +
  coord_flip() +
  theme(legend.position="none") + 
  labs(title = "Race Mixing Among Egos w/ 3 Active Partners")

g3 + facet_grid(cols=vars(e.sex))

```

## Weighted 

this gets weirder b/c svy design package doesn't seem to handle NAs well, errors when trying to generate counts I think b/c SEs are calculated automatically and they fail 

```{r mixing-svy, echo=FALSE}
svy <- as_survey_design(dat, weights=e.weight, ids=ego)
parts1 <- svy %>% filter(e.deg.active==1) %>%
              group_by(e.sex, e.race, race) %>% 
              summarize(prop = survey_mean()) %>% 
              select(-prop_se) 

p <- ggplot(parts1, aes(x=e.race, y=race)) + 
  geom_point(color="blue", alpha=0.3, aes(size=prop)) +
  geom_text(aes(label=round(prop,2))) +
  scale_size_area(max_size = 30) +
  coord_flip() +
  theme(legend.position="none") + 
  labs(title = "Race Mixing Among Egos w/ 1 Active Partner")
  
p + facet_grid(cols=vars(e.sex))
```

turns out there are no 'other' egos w/ hispanic partners who have a degree of 2, removing 'other' egos from this graph 
```{r mixing2-svy, echo=FALSE}
parts2 <- svy %>% filter(e.deg.active==2 & (e.race %in% "b" | e.race %in% "h" | e.race %in% "w")) %>%
              group_by(e.sex, e.race, race) %>% 
              summarize(prop = survey_mean()) %>% 
              select(-prop_se) 

p2 <- ggplot(parts2, aes(x=e.race, y=race)) + 
  geom_point(color="blue", alpha=0.3, aes(size=prop)) +
  geom_text(aes(label=round(prop,2))) +
  scale_size_area(max_size = 30) +
  coord_flip() +
  theme(legend.position="none") + 
  labs(title = "Race Mixing Among Egos w/ 2 Active Partners")
  
p2 + facet_grid(cols=vars(e.sex))
```

