# Egos 

```{r full-data, echo=FALSE}
nsfg <- readRDS("~/nsfg_analysis/data/nsfg_complete.rds")

svy <- as_survey_design(nsfg, weights = weight, ids = CASEID)
```

The full dataset from 2006-2015 includes data from `r nsfg %>% count()` respondents. What follows are some basic unweighted & weighted demographic descriptives for respondents as well as other of sexual partnership variables of interest. 


## Sex 
no NAs
```{r full-sex, echo=FALSE}
sex <- nsfg %>% count(sex)

p <- sex %>% 
  plot_ly(x= ~sex, y= ~n, type='bar', name="raw", color = I("indianred3")) %>%
  layout(yaxis = list(title="", barmode='group'), 
         xaxis = list(title=""),
         title = "Counts - Raw Data")


sex_w <- svy %>% group_by(sex) %>% summarize(count = survey_total(vartype="ci"))
p2 <- sex_w %>% plot_ly(x=~sex, y=~count, type="bar", name="weighted", color = I("indianred4"),
                        error_y=list(type="data",
                           symmetric=FALSE,
                           array=c(sex_w$count_upp-sex_w$count),
                           arrayminus=c(sex_w$count-sex_w$count_low),
                           color="grey90")) %>% 
         layout(yaxis = list(title="", barmode='group'), 
         xaxis = list(title=""), 
         title = "Counts")

subplot(p,p2)
```

## Age 
No NAs
```{r full-age, echo=FALSE}
ageF <- nsfg %>% filter(sex %in% "F") %>% count(age)
ageM <- nsfg %>% filter(sex %in% "M") %>% count(age)
age <- cbind(ageF, ageM[,2])
colnames(age)[2:3] <- c("Females", "Males") 

p <- age %>% 
  plot_ly(x= ~age, y= ~Females, type='bar', name="Females") %>%
  add_trace(x=~age, y=~Males, name="Males") %>%
  layout(yaxis = list(title="", barmode='group'), 
         xaxis = list(title=""),
         title = "Counts - Raw Data")
p

# weighted 
p2 <- svy %>% group_by(sex, age) %>% summarize(count=survey_total()) 
p2 <- dcast(p2, age~sex, value.var="count")
  
p2 %>% 
  plot_ly(x=~age, y=~`F`, type="bar", name="Females") %>%
  add_trace(x=~age, y=~M, type="bar", name="Males") %>%
    layout(yaxis = list(title="", barmode='group'), 
         xaxis = list(title=""),
         title = "Counts - Weighted Data")
```

## Race
### Proportion, All
No NAs
```{r full-race, echo=FALSE}
#### pie charts ####
# of full race dist - not by sex
r1 <- nsfg %>% count(race) %>% mutate(prop = n/sum(n))
r2 <- svy %>% group_by(race) %>% summarize(n=survey_total()) %>% mutate(prop = n/sum(n))

props <- rbind(r1[,3], r2[,4]) 
props$id <- c(rep("raw",4), rep("weighted", 4))
props$race <- c(rep(c("black", "hispanic", "other", "white"),2))

ggplot(props, aes(x = id, y = prop, fill = race)) + geom_bar(stat = 'identity') + 
  scale_fill_manual(values=c('#1f77b4',  '#ff7f0e', '#d62728', '#2ca02c'))
  
```

### Counts, By Sex
```{r racesex, echo = FALSE} 
#### by sex  ####
raceF <- nsfg %>% filter(sex %in% "F") %>% count(race)
raceM <- nsfg %>% filter(sex %in% "M") %>% count(race)

race <- cbind(raceF, raceM[,2])
colnames(race)[2:3] <- c("Females", "Males") 

race %>% 
  plot_ly(x= ~race, y= ~Females, type='bar', name="Females") %>%
  add_trace(x=~race, y=~Males, name="Males") %>%
  layout(yaxis = list(title="", barmode='group'), 
         xaxis = list(title="Raw Data"))

# weighted
rw <- svy %>% group_by(sex, race) %>% summarize(n=survey_total())
rw <- dcast(rw, race~sex, value.var = "n")
rw %>% plot_ly(x=~race, y=~`F`, type = 'bar', name="Females") %>% 
        add_trace(x=~race, y=~M, name="Males") %>%
        layout(xaxis = list(title="Weighted"))
```

## Age Category & Race
```{r full-agerace, echo=FALSE}
age.race <- nsfg %>% count(agecat, race)
arboth <- dcast(age.race, agecat~race, value.var = "n")

p <- arboth %>% 
  plot_ly(x=~agecat, y= ~b, type='bar', name="Black") %>%
  add_trace(x=~agecat, y=~h, name="Hispanic") %>%
  add_trace(x=~agecat, y=~w, name="White") %>%
  add_trace(x=~agecat, y=~o, name="Other") %>%
  layout(yaxis = list(title="", barmode='group'), 
         xaxis = list(title=""), title="Counts- Raw Data")
p

## weighted
arw <- svy %>% group_by(agecat, race) %>% summarize(count=survey_total())
arw <- dcast(arw, agecat~race, value.var = "count")

arw %>% plot_ly(x=~agecat, y=~b, type='bar', name="Black") %>%
          add_trace(x=~agecat, y=~h, name="Hispanic") %>%
          add_trace(x=~agecat, y=~w, name="White") %>%
          add_trace(x=~agecat, y=~o, name="Other") %>%
          layout(yaxis = list(title="", barmode='group'), 
               xaxis = list(title=""), title="Counts - Weighted Data")
```

## Counts (raw)

```{r pops, echo=FALSE}
age.raceF <- nsfg %>% filter(sex %in% "F") %>% count(agecat, race)
age.raceM <- nsfg %>% filter(sex %in% "M") %>% count(agecat, race)
Fem <- dcast(age.raceF, agecat~race, value.var = "n")
Male <- dcast(age.raceM, agecat~race, value.var = "n")
```
### Females
```{r fempop, echo=FALSE}
kable(Fem, title="Females") %>% kable_styling(position="center", full_width = F)
```
### Males
```{r malepop, echo=FALSE}
kable(Male, title="Males") %>% kable_styling(position="center", full_width = F)
```
## needs work - Sex Group FEMALES?? 
Of those sexually active in the last year: 
```{r sexpref, echo=FALSE}

grp <- nsfg %>% count(sexgrp.yr)

grp %>% plot_ly(x=~sexgrp.yr, y=~n, type='bar') %>%
    layout(yaxis = list(title=""), 
         xaxis = list(title=""))
```

