# Behavior {#epi}

__Behavior Among Those Sexually Active (in at least 1 active partnership)__

```{r setup-epi-knitr, echo=FALSE}
#suppressMessages(library(tidyverse))
#suppressMessages(library(srvyr))
#suppressMessages(library(kableExtra))
#suppressMessages(library(reshape2))
#suppressMessages(library(plotly))

# full dataset (egos < 30)
nsfg <- readRDS("~/NSFG_DATA/Full/nsfg_complete.rds")
svy <- as_survey_design(nsfg, weights = weight, ids = ego)

# long alters list 
alters <- readRDS("~/NSFG_DATA/Objects/altersegos_survdat.rds")
svy_alters <- as_survey_design(alters, weights = weight, ids = ego)

```
- sex freq
- condom use
(by age cat, sex, race, partnership type? degree?)   


## Frequency of Sex In Last Week
(for those inds in at least 1 partnership on the day of interview)
("sex4wk" / 4)

```{r sex-freq, echo=FALSE}
sexF <- svy %>% 
        filter(deg.main >=1, sex4wk <995, sex %in% "F") %>% 
        mutate(sex1wk = sex4wk/4) %>%
        group_by(agecat, race) %>%
        summarize(mean = survey_mean(sex1wk)) %>%
        mutate(mean = round(mean, 2)) %>%
        dcast(agecat~race, value.var = "mean")
sexM <- svy %>% 
        filter(deg.main >=1, sex4wk <995, sex %in% "M") %>% 
        mutate(sex1wk = sex4wk/4) %>%
        group_by(agecat, race) %>%
        summarize(mean = survey_mean(sex1wk)) %>%
        mutate(mean = round(mean, 2)) %>%
        dcast(agecat~race, value.var = "mean")

kable(sexF, caption = "Females") %>% 
  kable_styling(full_width = F, bootstrap_options = "striped", position="float_left")

kable(sexM, caption = "Males") %>% 
  kable_styling(full_width = F, bootstrap_options = "striped", position="right")

```

## Condom Use 
### Probability of Condom Use in Last Week  
("sex4wk" / "cond4wk")  

```{r condoms, echo=FALSE}
condF <- svy %>% 
        filter(deg.main >=1, sex4wk <995, sex %in% "F") %>% 
        group_by(agecat, race) %>%
        summarize(mean = survey_mean(p.cond)) %>%
        mutate(mean = round(mean, 2)) %>%
        dcast(agecat ~ race, value.var = "mean")

condM <- svy %>% 
        filter(deg.main >=1, sex4wk <995, sex %in% "M") %>% 
        group_by(agecat, race) %>%
        summarize(mean = survey_mean(p.cond)) %>%
        mutate(mean = round(mean, 2)) %>%
        dcast(agecat ~ race, value.var = "mean")
  

kable(condF, format='html', caption = "Females") %>% 
  kable_styling(full_width = F, bootstrap_options = "striped", position="float_left")

kable(condM, format='html', caption = "Males") %>% 
  kable_styling(full_width = F, bootstrap_options = "striped", position = "right")
```


### Condom use at last sex with partner - females only  

```{r conds-last, echo=FALSE, fig.align="center"}
ls1 <- svy_alters %>% filter(active ==1, e.sex %in% "F", optype <6, condls < 6) %>%
        mutate(optype = factor(optype, labels = c("CurSpouse", "CurCohab","FmrSpouse", "FmrCohab", "Other"))) %>%
        mutate(optype = as.factor(optype), condls = as.factor(condls)) %>%
        group_by(optype, condls) %>%
        summarize(n=survey_total()) %>% 
        group_by(optype) %>%
        mutate(prop = n/sum(n)) %>%
        dcast(optype~condls, value.var = "prop") %>%
        plot_ly(x=~optype, y=~`1`, type ='bar') %>%
        layout(yaxis = list(title="% Used A Condom", range=c(0,0.6)), 
               showlegend=FALSE)

ls2 <- svy_alters %>% filter(active==1, e.sex %in% "F", optype <6, condls < 6) %>%
        mutate(optype = as.factor(e.agecat), condls = as.factor(condls)) %>%
        group_by(e.agecat, condls) %>%
        summarize(n=survey_total()) %>% 
        group_by(e.agecat) %>%
        mutate(prop = n/sum(n)) %>%
        dcast(e.agecat~condls, value.var = "prop") %>%
        plot_ly(x=~e.agecat, y=~`1`, type ='bar') %>%
        layout(yaxis = list(title="% Used A Condom", range=c(0,0.6)),
               showlegend=FALSE)
subplot(ls1,ls2, margin = 0.05)
```


### Condom use at first sex with partner
#### By Sex & Partnership Type
```{r conds-first, echo=FALSE, fig.align="center"}
fs1 <- svy_alters %>% filter(active==1, e.sex %in% "F", optype <6, condfs <6) %>%
          mutate(optype = factor(optype, labels = c("CurSpouse", "CurCohab","FmrSpouse", "FmrCohab", "Other"))) %>%
        mutate(optype = as.factor(optype), condfs = as.factor(condfs)) %>%
        group_by(optype, condfs) %>%
        summarize(n=survey_total()) %>% 
        group_by(optype) %>%
        mutate(prop = n/sum(n)) %>%
        dcast(optype~condfs, value.var = "prop") %>%
        plot_ly(x=~optype, y=~`1`, type ='bar', name = "Female Egos", 
                text= c("Cur. Spouse", "Cur. Cohab", "Former Spouse", "Former Cohab", "Other")) %>%
        layout(yaxis = list(title="% Used A Condom", range=c(0,1)))

fs2 <- svy_alters %>% filter(active==1, e.sex %in% "M", optype <6, condfs <6) %>%
          mutate(optype = factor(optype, labels = c("CurSpouse", "CurCohab","FmrSpouse", "FmrCohab", "Other"))) %>%
        mutate(optype = as.factor(optype), condfs = as.factor(condfs)) %>%
        group_by(optype, condfs) %>%
        summarize(n=survey_total()) %>% 
        group_by(optype) %>%
        mutate(prop = n/sum(n)) %>%
        dcast(optype~condfs, value.var = "prop") %>%
        plot_ly(x=~optype, y=~`1`, type ='bar', name = "Male Egos", 
                text= c("Cur. Spouse", "Cur. Cohab", "Former Spouse", "Former Cohab", "Other")) %>%
        layout(yaxis = list(title="% Used A Condom", range=c(0,1)))

suppressWarnings(subplot(fs1,fs2, margin = 0.05))
```


#### All Egos By Partnership Type
```{r condsfirst-all, echo=FALSE, fig.align="center"}
svy_alters %>% filter(active==1, optype <6, condfs <6) %>%
          mutate(optype = factor(optype, labels = c("CurSpouse", "CurCohab","FmrSpouse", "FmrCohab", "Other"))) %>%
        mutate(optype = as.factor(optype), condfs = as.factor(condfs)) %>%
        group_by(optype, condfs) %>%
        summarize(n=survey_total()) %>% 
        group_by(optype) %>%
        mutate(prop = n/sum(n)) %>%
        dcast(optype~condfs, value.var = "prop") %>%
        plot_ly(x=~optype, y=~`1`, type ='bar', name = "Yes", 
                text= c("Cur. Spouse", "Cur. Cohab", "Former Spouse", "Former Cohab", "Other")) %>%
        layout(yaxis = list(title="% Used A Condom", range=c(0,1)), 
               title="All Egos")
```