# Network Factors - 15-29 {#net-young}


```{r ego-prep-young, echo=FALSE, message=FALSE}
library(ergm.ego)
library(tidyverse)
library(reshape2)
library(srvyr)
library(plotly)
library(kableExtra)

# import objects
alters <- readRDS("~/NSFG_DATA/Objects/alters_all_for_young_egos.rds")
alters_young <- readRDS("~/NSFG_DATA/Objects/alters_young_for_young_egos.rds")
egos <- readRDS("~/NSFG_DATA/Objects/egos_young.rds")

# make ego data objects
## egos < 30 and alters any age
dat <- egodata(egos=egos, alters=alters, egoWt = egos$weight, egoIDcol = "ego")
## egos < 30 and alters < 30
dat2 <- egodata(egos=egos, alters=alters_young, egoWt = egos$weight, egoIDcol = "ego")

# full dataset (egos < 30)
nsfg <- readRDS("~/NSFG_DATA/Objects/nsfg_young_egos.rds")
svy <- as_survey_design(nsfg, weights = weight, ids = ego)
```

__Mean Degree of egos < 30: `r round(nrow(dat$alters)/nrow(dat$egos), 3)`__  
__Mean Degree of egos & alters < 30: `r round(nrow(dat2$alters)/nrow(dat$egos), 3)`__


## Degree

### Histograms
#### All Alters Vs Alters < 30
```{r degdist-young, echo=FALSE, out.width="50%", out.height="50%", fig.show='hold'}

degreedist(dat); degreedist(dat2)

degreedist(dat, by="sex"); degreedist(dat2, by="sex")

degreedist(dat, by="race"); degreedist(dat2, by="race")

degreedist(dat, by="agecat"); degreedist(dat2, by="agecat")
```

### Bubble Plots  
##### egos < 30,  all corresponding alters 
```{r deg-bubble-young, echo=FALSE, out.width="50%", out.height="50%", fig.show='hold'}
# excluding the n=14 respondents who report 2 or 3 active married/cohab partners

bubF <- svy %>%
          filter(sex %in% "F", deg.cohab < 2) %>%
          mutate(deg.cohab = as.factor(deg.cohab), deg.pers = as.factor(deg.pers)) %>%
          group_by(deg.cohab, deg.pers, .drop = FALSE) %>%
          summarize(n = survey_total()) %>% 
          mutate(prop = n/sum(n)) %>%
          select(-n, -n_se)

bubF <- suppressMessages(melt(bubF))

bubM <- svy %>%
          filter(sex %in% "M", deg.cohab < 2) %>%
          mutate(deg.cohab = as.factor(deg.cohab), deg.pers = as.factor(deg.pers)) %>%
          group_by(deg.cohab, deg.pers, .drop = FALSE) %>%
          summarize(n = survey_total()) %>% 
          mutate(prop = n/sum(n)) %>%
          select(-n, -n_se)

bubM <- suppressMessages(melt(bubM))

My_Theme = theme(
  axis.title = element_text(size = 16),
  axis.text = element_text(size=16),
  title = element_text(size=16))

bubFplot <- ggplot(bubF, aes(deg.pers, deg.cohab)) +
              geom_point(color="green", alpha=0.3, aes(size=value)) +
              scale_size_area(max_size = 30) +
              geom_text(aes(label=round(value,3), size=0.04)) +
              theme(legend.position="none") + 
              labs(title="Females") +
              My_Theme

bubMplot <- ggplot(bubM, aes(deg.pers, deg.cohab)) +
              geom_point(color="green", alpha=0.3, aes(size=value)) +
              scale_size_area(max_size = 30) +
              geom_text(aes(label=round(value,3), size=0.04)) +
              theme(legend.position="none") + 
              labs(title="Males") +
              My_Theme

#grid.arrange(bubFplot, bubMplot, ncol=2)
bubFplot 

bubMplot
```

##### egos & alters < 30
```{r deg-bubble-young2, echo=FALSE, out.width="50%", out.height="50%", fig.show='hold'}
# excluding the n=14 respondents who report 2 or 3 active married/cohab partners

bubF <- svy %>%
          filter(sex %in% "F", deg.cohab < 2) %>%
          mutate(deg.cohab = as.factor(deg.cohab.y), deg.pers = as.factor(deg.pers.y)) %>%
          group_by(deg.cohab, deg.pers, .drop = FALSE) %>%
          summarize(n = survey_total()) %>% 
          mutate(prop = n/sum(n)) %>%
          select(-n, -n_se)

bubF <- suppressMessages(melt(bubF))

bubM <- svy %>%
          filter(sex %in% "M", deg.cohab < 2) %>%
          mutate(deg.cohab = as.factor(deg.cohab.y), deg.pers = as.factor(deg.pers.y)) %>%
          group_by(deg.cohab, deg.pers, .drop = FALSE) %>%
          summarize(n = survey_total()) %>% 
          mutate(prop = n/sum(n)) %>%
          select(-n, -n_se)

bubM <- suppressMessages(melt(bubM))

My_Theme = theme(
  axis.title = element_text(size = 16),
  axis.text = element_text(size=16),
  title = element_text(size=16))

bubFplot <- ggplot(bubF, aes(deg.pers, deg.cohab)) +
              geom_point(color="green", alpha=0.3, aes(size=value)) +
              scale_size_area(max_size = 30) +
              geom_text(aes(label=round(value,3), size=0.04)) +
              theme(legend.position="none") + 
              labs(title="Females") +
              My_Theme

bubMplot <- ggplot(bubM, aes(deg.pers, deg.cohab)) +
              geom_point(color="green", alpha=0.3, aes(size=value)) +
              scale_size_area(max_size = 30) +
              geom_text(aes(label=round(value,3), size=0.04)) +
              theme(legend.position="none") + 
              labs(title="Males") +
              My_Theme

#grid.arrange(bubFplot, bubMplot, ncol=2)
bubFplot 

bubMplot
```


## Age Mixing  

### Overall  
```{r age-mixing-young, echo=FALSE, out.width="50%", out.height="50%", fig.show='hold'}

agemix <- round(mixingmatrix(dat, "agecat", rowprob = T), 3)
am <- melt(agemix[c(2:4),])

am %>% ggplot(aes(ego, alter)) + 
  geom_point(color="blue", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Age Mixing") 

agemix2 <- round(mixingmatrix(dat2, "agecat", rowprob = T), 3)
am2 <- melt(agemix2)

am2 %>% ggplot(aes(ego, alter)) + 
  geom_point(color="blue", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Age Mixing (all under 30)") +
  My_Theme
```

### By Sex  

```{r age-mixing-young-sex, echo=FALSE, out.width="50%", out.height="50%", fig.show='hold'}

agemixF <- round(mixingmatrix(dat[which(dat$egos$sex %in% "F")], "agecat", rowprob = T), 3)
amF <- melt(agemixF[c(2:4),])

amF %>% ggplot(aes(ego, alter)) + 
  geom_point(color="blue", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Age Mixing, Females under 30")

agemixF2 <- round(mixingmatrix(dat2[which(dat2$egos$sex %in% "F")], "agecat", rowprob = T), 3)
amF2 <- melt(agemixF2) 

amF2 %>% ggplot(aes(ego, alter)) + 
  geom_point(color="blue", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Age Mixing, Females & Alters under 30") +
  My_Theme

agemixM <- round(mixingmatrix(dat[which(dat$egos$sex %in% "M")], "agecat", rowprob = T), 3)
amM <- melt(agemixM[c(2:4),])

amM %>% ggplot(aes(ego, alter)) + 
  geom_point(color="blue", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Age Mixing, Males under 30")

agemixM2 <- round(mixingmatrix(dat2[which(dat2$egos$sex %in% "M")], "agecat", rowprob = T), 3)
am2 <- melt(agemixM2)

am2 %>% ggplot(aes(ego, alter)) + 
  geom_point(color="blue", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Age Mixing, Males & Alters under 30") +
  My_Theme
```


### By Race  

```{r age-mixing-young-race, echo=FALSE, out.width="50%", out.height="50%", fig.show='hold'}
### Blacks 
agemixB <- round(mixingmatrix(dat[which(dat$egos$race %in% "b")], "agecat", rowprob = T), 3)
amB <- melt(agemixB[c(2:4),])

amB %>% ggplot(aes(ego, alter)) + 
  geom_point(color="blue", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Age Mixing, Blacks under 30")

agemixB2 <- round(mixingmatrix(dat2[which(dat$egos$race %in% "b")], "agecat", rowprob = T), 3)
amB2 <- melt(agemixB2) 

amB2 %>% ggplot(aes(ego, alter)) + 
  geom_point(color="blue", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Age Mixing, Blacks & Alters under 30") +
  My_Theme

### Hispanics
agemixH <- round(mixingmatrix(dat[which(dat$egos$race %in% "h")], "agecat", rowprob = T), 3)
amH <- melt(agemixH[c(2:4),])

amH %>% ggplot(aes(ego, alter)) + 
  geom_point(color="blue", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Age Mixing, Hispanics under 30")

agemixH2 <- round(mixingmatrix(dat2[which(dat$egos$race %in% "h")], "agecat", rowprob = T), 3)
amH2 <- melt(agemixH2)

amH2 %>% ggplot(aes(ego, alter)) + 
  geom_point(color="blue", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Age Mixing, Males & Alters under 30") +
  My_Theme

## Others
agemixO <- round(mixingmatrix(dat[which(dat$egos$race %in% "o")], "agecat", rowprob = T), 3)
amO <- melt(agemixO[c(2:4),])

amO %>% ggplot(aes(ego, alter)) + 
  geom_point(color="blue", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Age Mixing, Others under 30")

agemixO2 <- round(mixingmatrix(dat2[which(dat$egos$race %in% "o")], "agecat", rowprob = T), 3)
amO2 <- melt(agemixO2)

amO2 %>% ggplot(aes(ego, alter)) + 
  geom_point(color="blue", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Age Mixing, Others & Alters under 30") +
  My_Theme

## Whites 
agemixW <- round(mixingmatrix(dat[which(dat$egos$race %in% "w")], "agecat", rowprob = T), 3)
amW <- melt(agemixW[c(2:4),])

amW %>% ggplot(aes(ego, alter)) + 
  geom_point(color="blue", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Age Mixing, Whites under 30")

agemixW2 <- round(mixingmatrix(dat2[which(dat$egos$race %in% "w")], "agecat", rowprob = T), 3)
amW2 <- melt(agemixW2)

amW2 %>% ggplot(aes(ego, alter)) + 
  geom_point(color="blue", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Age Mixing, Whites & Alters under 30") +
  My_Theme
```

   
   
   
## Race Mixing 
### Overall  
```{r race-mixing-young, echo=FALSE, out.width="50%", out.height="50%", fig.show='hold'}
racemix <- round(mixingmatrix(dat, "race", rowprob = T), 3)
racemix <- racemix[c(1,2,5,6), c(1,2,5,6)]
rm <- melt(racemix) 

rm %>% ggplot(aes(ego, alter)) + 
  geom_point(color="indianred", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Race Mixing (many `other` partners missing race, excluded)") 

racemix2 <- round(mixingmatrix(dat2, "race", rowprob = T), 3)
racemix2 <- racemix2[c(1,2,5,6), c(1,2,5,6)]
rm2 <- melt(racemix2) 

rm2 %>% ggplot(aes(ego, alter)) + 
  geom_point(color="indianred", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Race Mixing (all under 30)") +
  My_Theme
```

### By Sex  

```{r race-mixing-young-2, echo=FALSE, out.width="50%", out.height="50%", fig.show='hold'}
racemixF <- round(mixingmatrix(dat[which(dat$egos$sex %in% "F")], "race", rowprob = T), 3)
racemixF <- racemixF[c(1,2,5,6), c(1,2,5,6)]
rmF <- melt(racemixF) 

rmF %>% ggplot(aes(ego, alter)) + 
  geom_point(color="indianred", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Race Mixing, Female Egos < 30") +
  My_Theme

racemixF2 <- round(mixingmatrix(dat2[which(dat2$egos$sex %in% "F")], "race", rowprob = T), 3)
racemixF2 <- racemixF2[c(1,2,5,6), c(1,2,5,6)]
rmF2 <- melt(racemixF2) 

rmF2 %>% ggplot(aes(ego, alter)) + 
  geom_point(color="indianred", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Race Mixing, Female Egos & Alters < 30") +
  My_Theme


racemixM <- round(mixingmatrix(dat[which(dat$egos$sex %in% "M")], "race", rowprob = T), 3)
racemixM <- racemixM[c(1,2,4,5), c(1,2,4,5)]
rmM <- melt(racemixM) 

rmM %>% ggplot(aes(ego, alter)) + 
  geom_point(color="indianred", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Race Mixing, Males Egos < 30") +
  My_Theme


racemixM2 <- round(mixingmatrix(dat2[which(dat2$egos$sex %in% "M")], "race", rowprob = T), 3)
racemixM2 <- racemixM2[c(1,2,4,5), c(1,2,4,5)]
rmM2 <- melt(racemixM2) 

rm2 %>% ggplot(aes(ego, alter)) + 
  geom_point(color="indianred", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Race Mixing, Males Egos & Alters < 30") +
  My_Theme
```

### By AgeCat
```{r race-mixing-young-3, echo=FALSE, out.width="50%", out.height="50%", fig.show='hold'}
### Egos 15-19
racemix15 <- round(mixingmatrix(dat[which(dat$egos$agecat %in% "15-19")], "race", rowprob = T), 3)
racemix15 <- racemix15[c(1,2,5,6), c(1,2,5,6)]
rm15 <- melt(racemix15) 

rm15 %>% ggplot(aes(ego, alter)) + 
  geom_point(color="indianred", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Race Mixing, Egos 15-19") +
  My_Theme

racemix15.2 <- round(mixingmatrix(dat2[which(dat2$egos$agecat %in% "15-19")], "race", rowprob = T), 3)
racemix15.2 <- racemix15.2[c(1,2,5,6), c(1,2,5,6)]
rm15.2<- melt(racemix15.2) 

rm15.2 %>% ggplot(aes(ego, alter)) + 
  geom_point(color="indianred", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Race Mixing, Egos 15-19 & Alters < 30") +
  My_Theme

#### Egos 20-24
racemix20 <- round(mixingmatrix(dat[which(dat$egos$agecat %in% "20-24")], "race", rowprob = T), 3)
racemix20 <- racemix20[c(1,2,5,6), c(1,2,5,6)]
rm20 <- melt(racemix20) 

rm20 %>% ggplot(aes(ego, alter)) + 
  geom_point(color="indianred", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Race Mixing, Egos 20-24") +
  My_Theme

racemix20.2 <- round(mixingmatrix(dat2[which(dat2$egos$agecat %in% "20-24")], "race", rowprob = T), 3)
racemix20.2 <- racemix20.2[c(1,2,5,6), c(1,2,5,6)]
rm20.2<- melt(racemix20.2) 

rm20.2 %>% ggplot(aes(ego, alter)) + 
  geom_point(color="indianred", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Race Mixing, Egos 15-19 & Alters < 30") +
  My_Theme

### 25-29
racemix25 <- round(mixingmatrix(dat[which(dat$egos$agecat %in% "25-29")], "race", rowprob = T), 3)
racemix25 <- racemix25[c(1,2,4,5), c(1,2,4,5)]
rm25 <- melt(racemix25) 

rm25 %>% ggplot(aes(ego, alter)) + 
  geom_point(color="indianred", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Race Mixing, Egos 25-29") +
  My_Theme

racemix25.2 <- round(mixingmatrix(dat2[which(dat2$egos$agecat %in% "25-29")], "race", rowprob = T), 3)
racemix25.2 <- racemix25.2[c(1,2,4,5), c(1,2,4,5)]
rm25.2<- melt(racemix25.2) 

rm25.2 %>% ggplot(aes(ego, alter)) + 
  geom_point(color="indianred", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.03)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Race Mixing, Egos 20-24 & Alters < 30") +
  My_Theme

```

### One-Times (all alters)
(mean one-times in last year by sex, race, agecat- inactive one-time partners)
```{r one-times, echo=FALSE}
inst_race <- svy %>% 
  filter(HADSEX==1) %>% 
  group_by(sex, race) %>% 
  summarize(mean = survey_mean(instyr)) %>% 
  mutate(mean = round(mean, 4)) %>%
  dcast(sex~race, value.var = "mean")

inst_age <- svy %>% 
  filter(HADSEX==1) %>% 
  group_by(sex, agecat) %>% 
  summarize(mean = survey_mean(instyr)) %>% 
  mutate(mean = round(mean, 4)) %>%
  dcast(sex~agecat, value.var = "mean")

kable(inst_race, caption = "By Sex & Ego Race") %>% 
  kable_styling(full_width = F, bootstrap_options = "striped", position="left")
```



```{r inst-age, echo=FALSE}
kable(inst_age, caption = "By Sex & Age Cat") %>% 
  kable_styling(full_width = F, bootstrap_options = "striped", position="left")
```