# Network Factors - ALL NSFG {#net-full}

## Degree
### Histograms

```{r ego-prep, echo=FALSE, message=FALSE}
library(ergm.ego)
library(tidyverse)
library(reshape2)
library(srvyr)
library(plotly)
library(kableExtra)


##### import data for egodata ########
egodat <- readRDS("~/NSFG_DATA/Objects/fullEgodata.rds")

##### make ego data objects #########
## all actives
dat <- egodata(egos=egodat$egos, alters=egodat$altersAllActive, egoWt = egodat$egos$weight, egoIDcol = "ego")
## mar/coh
datMarcoh <- egodata(egos=egodat$egos, alters=egodat$altersMarCoh, egoWt = egodat$egos$weight, egoIDcol = "ego")
## others 
datOther <- egodata(egos=egodat$egos, alters=egodat$altersOther, egoWt = egodat$egos$weight, egoIDcol = "ego")
## one times 
datOTs <- egodata(egos=egodat$egos, alters=egodat$altersOnce, egoWt = egodat$egos$weight, egoIDcol = "ego")

# full data
nsfg <- readRDS("~/NSFG_DATA/Full/nsfg_complete.rds")
svy <- as_survey_design(nsfg, weights = weight, ids = ego)
```

__Mean Degree: `r round(nrow(dat$alters)/nrow(dat$egos), 3)`__

```{r degdist, echo=FALSE, out.width="50%", out.height="50%", fig.show='hold'}
meandeg <- nrow(dat$alters)/nrow(dat$egos)

degreedist(dat)

degreedist(dat, by="sex")

degreedist(dat, by="race")

degreedist(dat, by="agecat")
```

### Bubble Plots  

```{r deg-bubble, echo=FALSE, out.width="50%", out.height="50%", fig.show='hold'}
# excluding the n=14 respondents who report 2 or 3 active married/cohab partners

bubF <- svy %>%
          filter(sex %in% "F", deg.cohab < 2) %>%
          mutate(deg.cohab = as.factor(deg.cohab), deg.pers = as.factor(deg.pers)) %>%
          group_by(deg.cohab, deg.pers, .drop = FALSE) %>%
          summarize(n = survey_total()) %>% 
          mutate(prop = n/sum(n)) %>%
          select(-n, -n_se)

bubF <- suppressMessages(melt(bubF))

bubM <- svy %>%
          filter(sex %in% "M", deg.cohab < 2) %>%
          mutate(deg.cohab = as.factor(deg.cohab), deg.pers = as.factor(deg.pers)) %>%
          group_by(deg.cohab, deg.pers, .drop = FALSE) %>%
          summarize(n = survey_total()) %>% 
          mutate(prop = n/sum(n)) %>%
          select(-n, -n_se)

bubM <- suppressMessages(melt(bubM))

My_Theme = theme(
  axis.title = element_text(size = 16),
  axis.text = element_text(size=16),
  title = element_text(size=16))

bubFplot <- ggplot(bubF, aes(deg.pers, deg.cohab)) +
              geom_point(color="green", alpha=0.3, aes(size=value)) +
              scale_size_area(max_size = 30) +
              geom_text(aes(label=round(value,3), size=0.04)) +
              theme(legend.position="none") + 
              labs(title="Females") +
              My_Theme

bubMplot <- ggplot(bubM, aes(deg.pers, deg.cohab)) +
              geom_point(color="green", alpha=0.3, aes(size=value)) +
              scale_size_area(max_size = 30) +
              geom_text(aes(label=round(value,3), size=0.04)) +
              theme(legend.position="none") + 
              labs(title="Males") +
              My_Theme

#grid.arrange(bubFplot, bubMplot, ncol=2)
bubFplot 

bubMplot
```

## Age Mixing 

### Overall 
```{r age-mixing, echo=FALSE, fig.align="center"}
agemixF <- round(mixingmatrix(dat[dat$egos$sex %in% "F"], "agecat", rowprob = T), 3)
agemixF <- agemixF[2:7,]
amF <- melt(agemixF)
amF$sex <- "F"

agemixM <- round(mixingmatrix(dat[dat$egos$sex %in% "M"], "agecat", rowprob = T), 3)
agemixM <- agemixM[2:7,]
amM <- melt(agemixM)
amM$sex <- "M"

am <-  rbind(amF, amM)

am %>% ggplot(aes(ego, alter)) + 
  geom_point(color="blue", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.01)) +
  theme(legend.position="none",
        axis.text.x = element_text(angle=45)) +
  coord_flip() +
  labs(title = "Age Mixing - All Active Rels") +
  facet_wrap(~sex, ncol = 2)
```

### Marriages / Cohabs

```{r age-marcoh, echo=FALSE, fig.align="center"}
agemixF <- round(mixingmatrix(datMarcoh[datMarcoh$egos$sex %in% "F"], "agecat", rowprob = T), 3)
agemixF <- agemixF[2:7,]
amF <- melt(agemixF)
amF$sex <- "F"

agemixM <- round(mixingmatrix(datMarcoh[datMarcoh$egos$sex %in% "M"], "agecat", rowprob = T), 3)
agemixM <- agemixM[2:7,]
amM <- melt(agemixM)
amM$sex <- "M"

am <-  rbind(amF, amM)

am %>% ggplot(aes(ego, alter)) + 
  geom_point(color="blue", alpha=0.2, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.015)) +
  theme(legend.position="none",
        axis.text.x = element_text(angle=45)) +
  coord_flip() +
  labs(title = "Age Mixing - All Active Mar/Cohs") +
  facet_wrap(~sex, ncol = 2)
```

### Others 
```{r age-other, echo=FALSE, fig.align="center"}
agemixF <- round(mixingmatrix(datOther[datOther$egos$sex %in% "F"], "agecat", rowprob = T), 3)
agemixF <- agemixF[2:7,]
amF <- melt(agemixF)
amF$sex <- "F"

agemixM <- round(mixingmatrix(datOther[datOther$egos$sex %in% "M"], "agecat", rowprob = T), 3)
agemixM <- agemixM[2:7,]
amM <- melt(agemixM)
amM$sex <- "M"

am <-  rbind(amF, amM)

am %>% ggplot(aes(ego, alter)) + 
  geom_point(color="blue", alpha=0.2, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.015)) +
  theme(legend.position="none",
        axis.text.x = element_text(angle=45)) +
  coord_flip() +
  labs(title = "Age Mixing - All Active Others") +
  facet_wrap(~sex, ncol = 2)
```

### One-Times

this needs some work 

```{r agemixing-onetimes, echo=FALSE}
agemixF <- round(mixingmatrix(datOTs[datOTs$egos$sex %in% "F"], "agecat", rowprob = T), 3)
agemixF <- agemixF[1:6,]
amF <- melt(agemixF)
amF$sex <- "F"

agemixM <- round(mixingmatrix(datOTs[datOTs$egos$sex %in% "M"], "agecat", rowprob = T), 3)
agemixM <- agemixM[1:6,]
amM <- melt(agemixM)
amM$sex <- "M"

am <-  rbind(amF, amM)

am %>% ggplot(aes(ego, alter)) + 
  geom_point(color="blue", alpha=0.2, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.015)) +
  theme(legend.position="none",
        axis.text.x = element_text(angle=45)) +
  coord_flip() +
  labs(title = "Age Mixing - One-Times") +
  facet_wrap(~sex, ncol = 2)
```

## Race Mixing 
```{r race-mixing-all, echo=FALSE, fig.align="center"}
racemixF <- round(mixingmatrix(dat[dat$egos$sex %in% "F"], "race", rowprob = T), 3)
rmF <- melt(racemixF)
rmF$sex <- "F"

racemixM <- round(mixingmatrix(dat[dat$egos$sex %in% "M"], "race", rowprob = T), 3)
rmM <- melt(racemixM)
rmM$sex <- "M"

rm <-  rbind(rmF, rmM)

rm %>% ggplot(aes(ego, alter)) + 
  geom_point(color="springgreen3", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.015)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Race Mixing - All Active Rels") +
  facet_wrap(~sex, ncol = 2)
```

### Mar/Cohs
```{r race-mixing-marcoh, echo=FALSE, fig.align="center"}
racemixF <- round(mixingmatrix(datMarcoh[datMarcoh$egos$sex %in% "F"], "race", rowprob = T), 3)
rmF <- melt(racemixF)
rmF$sex <- "F"

racemixM <- round(mixingmatrix(datMarcoh[datMarcoh$egos$sex %in% "M"], "race", rowprob = T), 3)
rmM <- melt(racemixM)
rmM$sex <- "M"

rm <-  rbind(rmF, rmM)

rm %>% ggplot(aes(ego, alter)) + 
  geom_point(color="springgreen3", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.015)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Race Mixing - All Active Mar/Coh") +
  facet_wrap(~sex, ncol = 2)
```

### Other
```{r race-mixing-other, echo=FALSE, fig.align="center"}
racemixF <- round(mixingmatrix(datOther[datOther$egos$sex %in% "F"], "race", rowprob = T), 3)
rmF <- melt(racemixF)
rmF$sex <- "F"

racemixM <- round(mixingmatrix(datOther[datOther$egos$sex %in% "M"], "race", rowprob = T), 3)
rmM <- melt(racemixM)
rmM$sex <- "M"

rm <-  rbind(rmF, rmM)

rm %>% ggplot(aes(ego, alter)) + 
  geom_point(color="springgreen3", alpha=0.3, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.015)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Race Mixing - All Active Others") +
  facet_wrap(~sex, ncol = 2)
```

### One-Times 

one-times based on Male egos reports of partner race if one-time partner was most recent 
female one-times partner race based on male reports 

```{r racemix-onetimes, echo=FALSE}
racemixM <- round(mixingmatrix(datOTs[datOTs$egos$sex %in% "M"], "race", rowprob = T), 3)
rmM <- melt(racemixM)
rmM$sex <- "M"

racemixMcounts<- mixingmatrix(datOTs[datOTs$egos$sex %in% "M"], "race")
racemixFcounts <- t(racemixMcounts)
racemixF <- round(racemixFcounts/rowSums(racemixFcounts),3) 
rmF <- melt(racemixF)
colnames(rmF) <- c("ego", "alter", "value")
rmF$sex <- "F"

rm <-  rbind(rmF, rmM)

rm %>% ggplot(aes(ego, alter)) + 
  geom_point(color="blue", alpha=0.2, aes(size=value)) +
  scale_size_area(max_size = 30) +
  geom_text(aes(label=round(value,2), size=0.015)) +
  theme(legend.position="none") +
  coord_flip() +
  labs(title = "Race Mixing - One-Times") +
  facet_wrap(~sex, ncol = 2)
```

## New Partnerships in Last Year

```{r new-partners, echo=FALSE}
# what proportion of all reported partnerships were new this year? 
prop <- nrow(dat$alters[which(dat$alters$dfs <= 12),])/nrow(dat$alters)
```

Of all active alters, `r round(prop,3)*100`% began within the last year. 

```{r new-partners-byage, echo=FALSE}
newParts <- function(dat, range){
  egos<- dat$egos$ego[which(dat$egos$agecat %in% range)]
  alters <- dat$alters[which(dat$alters$ego %in% egos),]
  prop <- nrow(alters[alters$dfs <= 12,])/nrow(alters)
  return(round(prop,3))
}

parts <- rbind(c("15-19","20-24","25-29", "30-34", "35-39", "40-44"),
              c(newParts(dat, "15-19"), newParts(dat, "20-24"), newParts(dat, "25-29"), newParts(dat, "30-34"), newParts(dat, "35-39"), newParts(dat, "40-44")))
rownames(parts) <- c("Ages", "Prop")
kable(parts)
```

## One-Times

(mean one-times in last year by sex, race, agecat- inactive one-time partners)
```{r one-times, echo=FALSE}
inst_race <- svy %>% 
  filter(HADSEX==1) %>% 
  group_by(sex, race) %>% 
  summarize(mean = survey_mean(instyr)) %>% 
  mutate(mean = round(mean, 4)) %>%
  dcast(sex~race, value.var = "mean")

inst_age <- svy %>% 
  filter(HADSEX==1) %>%
  group_by(sex, agecat) %>% 
  summarize(mean = survey_mean(instyr)) %>% 
  mutate(mean = round(mean, 4)) %>%
  dcast(sex~agecat, value.var = "mean")

kable(inst_race, caption = "By Sex & Ego Race") %>% 
  kable_styling(full_width = F, bootstrap_options = "striped", position="left")
```

```{r inst-age, echo=FALSE}
kable(inst_age, caption = "By Sex & Age Cat") %>% 
  kable_styling(full_width = F, bootstrap_options = "striped", position="left")
```