# Partnerships {#parts}


- partnership types 
- types by active / inactive 
- who had same-sex contact last year? / ever? 

## Sexual Activty in Last Year
### Females 
```{r activity-graph-females, echo=FALSE}
actF <- svy %>% filter(sex %in% "F") %>%
                mutate(act = ifelse(HADSEX == 2, "1", 
                            ifelse(PARTS1YR == 0, "2", 
                            ifelse(PARTS1YR > 0, "3", NA)))) %>%
                group_by(agecat, act) %>% 
                summarize(n=survey_total()) %>%
                group_by(agecat) %>%
                mutate(prop = n/sum(n))
actF <- dcast(actF, agecat~act, value.var = "prop")



actF %>% plot_ly(x=~agecat, y=~`1`, type='bar', name='Never Sex') %>%
          add_trace(y=~`2`, type='bar', name = "Inactive") %>%
          add_trace(y=~`3`, type='bar', name = "Active") %>%
          layout(title = "Sexual Activity by Age, Females", 
                 yaxis = list(title="Proportion"),
                 xaxis = list(title="Age Category"))


```
### Males
```{r sexual-activity-males, echo=FALSE}
actM <- svy %>% filter(sex %in% "M") %>%
                mutate(act = ifelse(HADSEX == 2, "1", 
                            ifelse(PARTS1YR == 0, "2", 
                            ifelse(PARTS1YR > 0, "3", NA)))) %>%
                group_by(agecat, act) %>% 
                summarize(n=survey_total()) %>%
                group_by(agecat) %>%
                mutate(prop = n/sum(n))
actM <- dcast(actM, agecat~act, value.var = "prop")

actM %>% plot_ly(x=~agecat, y=~`1`, type='bar', name='Never Sex') %>%
          add_trace(y=~`2`, type='bar', name = "Inactive") %>%
          add_trace(y=~`3`, type='bar', name = "Active") %>%
          layout(title = "Sexual Activity by Age, Males", 
                 yaxis = list(title="Proportion"),
                 xaxis = list(title="Age Category"))
```

```{r load-altersedgelist, echo=FALSE}
dat <- readRDS("~/NSFG_DATA/objects/alters_egos.rds")
# limit to partnership in last year 
dat_subset <- dat %>% filter(active < 9)
datsvy <- as_survey_design(dat, weights = e.weight, ids = ego)
datsvy_sub <- as_survey_design(dat_subset, weights = e.weight, ids = ego)
```
## Partnerships in Last Year 

```{r parts1yr, echo=FALSE}
partsyr <- suppressWarnings(svy %>% 
                              filter(PARTS1YR >= 0) %>% 
                              group_by(PARTS1YR) %>% 
                              summarize(n=survey_total()))

partsyr$PARTS1YR <- as.numeric(partsyr$PARTS1YR)
partsyr <- partsyr[order(partsyr$PARTS1YR),] 

partsyr %>% plot_ly(x=~PARTS1YR, y=~n, type="bar") %>%
            layout(title="parts. in last year among those who have had sex")

datatable(round(partsyr, 0))
```

```{r prop4, echo=FALSE, results='hide'}
props <- suppressWarnings(svy %>% 
                            mutate(parts_prop = ifelse(as.numeric(vsnpyr) < 4, 1, 
                            ifelse(as.numeric(vsnpyr) >= 4, 2, NA))) %>%
                            group_by(parts_prop) %>% 
                            summarize(n=survey_total()) %>%
                            mutate(prop = n/sum(n)))

propF <- suppressWarnings(svy %>% 
                            filter(sex %in% 'F') %>%
                            mutate(parts_prop = ifelse(as.numeric(vsnpyr) < 4, 1, 
                            ifelse(as.numeric(vsnpyr) >= 4, 2, NA))) %>%
                            group_by(parts_prop) %>% 
                            summarize(n=survey_total()) %>%
                            mutate(prop = n/sum(n)))

propM <- suppressWarnings(svy %>% 
                            filter(sex %in% 'M') %>%
                            mutate(parts_prop = ifelse(as.numeric(vsnpyr) < 4, 1, 
                            ifelse(as.numeric(vsnpyr) >= 4, 2, NA))) %>%
                            group_by(parts_prop) %>% 
                            summarize(n=survey_total()) %>%
                            mutate(prop = n/sum(n)))
```

Only `r props[2,4]*100`% of respondents report more than 3 partnership in the last year.  
`r propF[2,4]*100`% of Females  
`r propM[2,4]*100`% of Males  


## Active / Inactive 
(of the up to 3 detailed partnerships reported in the last year)

```{r weightedactives, echo=FALSE}
a1w <- suppressWarnings(datsvy_sub %>% 
  filter(active < 2) %>% 
  group_by(e.sex, active) %>% 
  summarize(count=survey_total()))

a1w <- dcast(a1w, active~e.sex, value.var = "count")
a1w$active <- c("0-Inactive", "1-Active")

a1p <- a1w %>% plot_ly(x=~active, y=~`F`, type='bar', name="Females") %>%
  add_trace(x=~active, y=~M, name="Males") %>%
  layout(title="Partnerships in Last Year", 
         yaxis = list(title = "Counts"),
         xaxis = list(title = ""))

a1p
``` 



## other ego vars
rel types ('optype')
sex group by age / have had ss contact by age 




## needs work - Sex Group FEMALES?? 
Of those sexually active in the last year: 
```{r sexpref, echo=FALSE}

grp <- nsfg %>% count(sexgrp.yr)

grp %>% plot_ly(x=~sexgrp.yr, y=~n, type='bar') %>%
    layout(yaxis = list(title=""), 
         xaxis = list(title=""))
```


